Project-II by Group Sydney

### 2015-12-21 Projet II/report/paper.tex
advance report and clean code

\begin{abstract}
This report provides a summary of the project two of the PCML class. The project consists of doing binary and multi-class classification on the same set of images. For both methods, we have observed that the given CNN features were more accurate than the HOG one. In binary classification, all methods were approximately equivalent except the Adaboosting and the SVM one, but we choose the SVM method as our final model. For the multi-class classification, neural networks and SVM were the best methods but neural networks has less variance in this case and is considered as our final model. The best way of improving our models would be to find a way to combine HOG features with the CNN one.

@@

Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 42273$ ($5408$ for the HOG and $36865$ for the CNN) and is the concatenation of $\mathbf{x}_n^{hog}$ with $\mathbf{x}_n^{cnn}$
The output ($\mathbf{y}$) represents the classification of the images. For the binary classification, the label $1$  represents a car, a horse or a plane and the label $2$ anything else. For the multi-class classification, the label $1$ represents a car, $2$2 a horse, $3$ a plane and $4$ anything else.

@@

We first have plot the labels distribution for the binary and multi-class cases. For the binary case, we have $60.3\%$ for the label $1$ and $39.8\%$ for the second one. We have observed that there is a small imbalance between the classes. For the multi-classes case, we obtain $16.07\%$, $19.37\%$, $24.87\%$ and $39.70\%$ for the labels $1$, $2$, $3$ and $4$. This time, the imbalance is more important, especially with the label $4$ and we do have to take care about it during the training.

We also wanted to analyze the distribution of our features, and have observed that for the features HOG, each of them has a has similar distribution. The values are essentially in the range $[0, 0.2]$. We are also interested about the correlation between the input and output variables. We have observed that the correlation is in the range $[-0.24 ;0.30]$, and so can conclude that features seem not to have a direct correlation between them. However, a combination of features can have a high correlation with the output but we couldn't find such combination. Finally, $\mathbf{X}_{hog}$ is rank-deficient with a rank of $3467$ instead of $5408$.

@@

This section is a description of the techniques we have used in the project. It describes mainly how it works, why we thought it could be useful in this project.

\begin{figure}[!ht]
\center
\subfigure[Plot of the distortion $J$ to $M$ where $M = D$. The distortion decreases quickly.]{\includegraphics[width=2.7in]{figures/distortionHOG.jpg} \label{fig:distHOG}}
\hfill
\subfigure[Plot of the distortion $J$ to $M$ where $M = N$ because there are too many features and so, the distortion doesn't go to $0$ and decreases slowly.]{\includegraphics[width=2.7in]{figures/distortionCNN.jpg}\label{fig:distCNN}}
\hfill
\caption{}
\end{figure}

@@

As we have seen in the section "Data visualization and analysis", we are faced to the $D > N$ problem mainly with the CNN features and also the HOG features depending on how we se the $k$ for the k-fold cross validation. One way to solve this problem is to use PCA.  Basically, PCA works as follow : we compute the mean and the covariance matrix $S = \frac{1}{N}\sum_{n=1}^{N}(x_n-\bar{x})(x_n-\bar{x})^T$ of our data and then compute the $M$ eigenvectors of $S$ corresponding to the $M$ largest eigenvalues. $M$ is the number of features we have to take into account in order to minimize the reconstruction. Finally, the approximation of a data sample is $\tilde{x_n} = \sum_{i=1}^M x_{ni} u_i$. However, we couldn't compute PCA using this method, because computing the full eigenvector decomposition a matrix $D \times D$ runs in  $O(D^3)$ and D is quite large. 
We have used an alternative proposed in [REF] which allows us to compute the same eigenvectors in $O(N^3)$, which is roughly at least $230$ times faster. The tricks is to express $S = \frac{1}{N}XX^T$ and finally using $\frac{1}{N}XX^Tv_i = \lambda_i v_i$, where $v_i=Xu_i$, we obtain an eigenvector equation for the $N \times N$ matrix $S$.

@@

The Figure \ref{fig:distHOG} shows the plot of $J$ for the HOG features and the CNN features. Figure \ref{fig:distCNN} shows the distortion of the CNN features, we have tried several values for $M$ and have conclude that $M_{CNN} = 150$ was a good value allowing us to compute faster results without loosing too much accuracy. The corresponding distortion might be very high compared to the HOG features, but the range of the values are not the same, CNN has values going larger than $0.2$. We tried some bigger values but we didn't get major improvements in terms of error. For the HOG features, the distortion measure seems more nice, we did the same experiments and finally set $M_{HOG}$ = $500$.

@@

This is the multi-layer perceptron (MLP) we have seen in class. The neural networks used in this project comes from the given DeepLearning matlab toolbox [REF]. It uses batch sizes in order to use less memory (take the first $n$ data samples and works on them, then the next $n$ etc). We have let the default parameter ($100$). The number of epochs is simply the number of time where all the training data do a forward and a backward pass. We have put this value to $20$, in order to have a little more accuracy without loosing a lot of time, it was a good compromise because larger epochs don't mean especially better accuracy. The number of input is $M$ and output is either $2$ or $4$, depending of the type of classification (binary or multi-class). For the number of hidden layer, we have let $1$ because we didn't see improvement by adding some more. However, the number of neurons in the hidden layer has been tested with k-fold cross validation and we have found $1000$ for the binary classification and $700$ for the multi-class case. We also have to set up the learning rate : the optimal value we have found is $3$ for the binary case and $2$ for the multi-class one. By default, the activation function used is the \textit{tanh} one. We didn't see so much different using the \textit{sigmoid}. Finally, the algorithm uses stochastic gradient descent using a momentum of $0.5$, which might help to converge faster.

@@

The idea of bootstrap aggregation is to average several decision trees. It will help to reduce the variance and avoid overfitting. It generates $m$ new training sets of the same size, by sampling uniformly and with replacement the original training set. Then we have $m$ decision trees which will grows independently with their own generated training sets. Finally, to predict the class it takes the average over all the trees. In order to find those parameters $m$ and the maximum depth of the tree, we have also used k-fold cross validation and the best values are $m = 200$ and maximum depth of $512$ for the binary classification and for the multi-class case $m = 500$ and maximum depth also of $512$. Having a high maximum depth might lead to overfit because we specialize to much the features split, but if it is too small, it might underfit because the features are not well split in order to identify classes. For the parameter $m$, it is better to have a higher number in order to have a large number of trees which can give a better accuracy of the prediction.

@@

The idea of random forests is similar than with bagging in a sense that we train several trees and take the average predictions. Like bagging, each trees have trained on a subset of the dataset. However, there is another aspect in random forest that bagging doesn't have: for each split, instead of searching among the overall features, we search only among a random subset of features. The size of this subset is a parameter $m$ as well as the maximum depth tree and the number of trees, like in bagging. Like bagging, random forests are useful because they help reducing the variance and the overfiting. But the more interesting ideas is to select only a subset of the features in the split process. Moreover, it is efficient against imbalanced dataset.

@@

The given Piotr matlab toolboox [REF] contains a method \textit{forestTrain} which takes the defined parameters and \textit{forestApply} for the prediction.

@@

Adaboosting is another way to average trees. If we observe more deeply the bagging approach, some trees might have a very high variance and some a very slow. Averaging all those trees decrease the variance and have a small bias. Boosting is the opposite : decrease the bias when averaging and have a small variance. In order to achieve this, each data sample has an assigned non-negative weight. After each iteration, if the data sample is misclassified, its weight is increased for the next iteration. So misclassified data samples will have more importance. The parameters are the same as in the bagging and behave in the same manner. The Piotr matlab toolboox [REF] contains a method \textit{adaBoostTrain} and \textit{adaBoostApply} for the prediction. We have only used it for the binary classification, because the implementation is not adapted for more classes. The parameters we obtain with k fold cross validation are $1024$ for the number of trees and $4$ for the maximum depth.

@@

In order to evaluate our different models, we have split our data in two sets of $80\%$ and $20\%$ for the training and the test sets. We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set. We learn only on the training set and then use the test set in order to estimate the error using the \textit{Balanced Error Rate}. This error compute the average percentage of misclassification per class. It allows us to avoid the case of the $90\%/10\%$ and having a good error. We have also used k-fold cross validation to tune the parameters up of our different models. We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set). We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.

@@

This section describes our models for the binary classification as well as the multi-class classification. We have essentially put our effort on the CNN features. The reason is because with only the HOG features, we couldn't get an error lower than $20\%$ for binary and multi-class classification. With the CNN features, we didn't have this problem. We have tried to take the whole matrix, only a subset of features with $M = 500$, different methods but nothing worked. Contrary to the CNN features, we think that the reason is because those features are too noisy : they came directly form the images. Some of them are rotated, inversed and so, the gradient is also transformed, which might be confusing for the learning. The CNN features are already computed from a convolutional neural networks which gives the features it has seen on the image with a specific weight. Because each sample has approximately $5\%$ of the overall features, they are fewer and are more specific than the HOG features.

@@

[1]    IMM2012-06284, R. B. Palm, Prediction as a candidate for learning deep hierarchical models of data, 2012.

@@

\end{document}

>>>>

\begin{abstract}
This report provides a summary of the project two of the PCML class. The project consists of doing binary and multi-class classification on the same set of images. For both methods, we have observed that the given CNN features were more accurate than the HOG one. In binary classification, all methods were approximately equivalent except the Adaboosting and the SVM one, but we choose the SVM method as our final model because it was more stable. For the multi-class classification, neural networks and SVM were the best methods but neural networks has less variance in this case and was considered as our final model. The best way of improving our models would be to find a way to combine HOG features with the CNN one.

@@

Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 42273$ ($5408$ for the HOG and $36865$ for the CNN) and is the concatenation of $\mathbf{x}_n^{hog}$ with $\mathbf{x}_n^{cnn}$.
The output ($\mathbf{y}$) represents the classification label of the images. For the binary classification, the label $1$  represents a car, a horse or a plane and the label $2$ anything else. For the multi-class classification, the label $1$ represents a car, $2$ a horse, $3$ a plane and $4$ anything else.

@@

We first have plot the labels distribution for the binary and multi-class cases. For the binary case, we have $60.3\%$ of the data for the label $1$ and $39.8\%$ for the second one. We have observed that there is a small imbalance between the classes. For the multi-classes case, we obtain $16.07\%$, $19.37\%$, $24.87\%$ and $39.70\%$ for the labels $1$, $2$, $3$ and $4$. This time, the imbalance is more important, especially with the label $4$ and we do have to take care about it during the training.

We also wanted to analyze the distribution of our features, and have observed that for the features HOG, each of them has a similar distribution. The values are essentially in the range $[0, 0.2]$. We are also interested about the correlation between the input and output variables. We have observed that the correlation is in the range $[-0.24 ;0.30]$, and so can conclude that features seem not to have a direct correlation between them. However, a combination of features can have a high correlation with the output but we couldn't find such combination. Finally, $\mathbf{X}_{hog}$ is rank-deficient with a rank of $3467$ instead of $5408$.

@@

This section is a description of the techniques we have used in the project. It describes mainly how it works, why we thought it could be useful in this project and the impact of their parameters.

\begin{figure}[!ht]
\center
\subfigure[Plot of the distortion $J$ to $M$ where $M = D$ for the HOG features. The distortion decreases quickly.]{\includegraphics[width=2.7in]{figures/distortionHOG.jpg} \label{fig:distHOG}}
\hfill
\subfigure[Plot of the distortion $J$ to $M$ for the CNN features, where $M = N$ because there are too many features and so, the distortion doesn't go to $0$ and decreases slowly.]{\includegraphics[width=2.7in]{figures/distortionCNN.jpg}\label{fig:distCNN}}
\hfill
\caption{}
\end{figure}

@@

As we have seen in the section "Data visualization and analysis", we are faced to the $D > N$ problem mainly with the CNN features and also the HOG features depending on how we se the $k$ for the k-fold cross validation. One way to solve this problem is to use PCA.  Basically, PCA works as follow : we compute the mean and the covariance matrix $S = \frac{1}{N}\sum_{n=1}^{N}(x_n-\bar{x})(x_n-\bar{x})^T$ of our data and then compute the $M$ eigenvectors of $S$ corresponding to the $M$ largest eigenvalues. $M$ is the number of features we have to take into account in order to minimize the reconstruction. Finally, the approximation of a data sample is $\tilde{x_n} = \sum_{i=1}^M x_{ni} u_i$. However, we couldn't compute PCA using this method, because computing the full eigenvector decomposition a matrix $D \times D$ runs in  $O(D^3)$ and D is quite large. 
We have used an alternative proposed in [3] which allows us to compute the same eigenvectors in $O(N^3)$, which is roughly at least $230$ times faster. The tricks is to express $S = \frac{1}{N}XX^T$ and finally using $\frac{1}{N}XX^Tv_i = \lambda_i v_i$, where $v_i=Xu_i$, we obtain an eigenvector equation for the $N \times N$ matrix $S$.

@@

The Figure \ref{fig:distHOG} shows the plot of $J$ for the HOG features. Figure \ref{fig:distCNN} shows the distortion of the CNN features, we have tried several values for $M$ and have conclude that $M_{CNN} = 150$ was a good value allowing us to compute faster results without loosing too much accuracy. The corresponding distortion might be very high compared to the HOG features, but the range of the values are not the same, CNN has values going larger than $0.2$. We tried some bigger values but we didn't get major improvements in terms of error. For the HOG features, the distortion measure seems more nice, we did the same experiments and finally set $M_{HOG}$ = $1000$.

@@

This is the multi-layer perceptron (MLP) we have seen in class. The neural networks used in this project comes from the given DeepLearning matlab toolbox [1]. It uses batch sizes in order to use less memory (take the first $n$ data samples and works on them, then the next $n$ etc). We have let the default parameter ($100$). The number of epochs is simply the number of time where all the training data do a forward and a backward pass. We have put this value to $20$, in order to have a little more accuracy without loosing a lot of time, it was a good compromise because larger epochs don't mean especially better accuracy. The number of input is $M$ and output is either $2$ or $4$, depending of the type of classification (binary or multi-class). For the number of hidden layer, we have let $1$ because we didn't see improvement by adding some more. However, if this number is to high, we might overfit and so have to be careful. Having this value to $1$ might not lead to overfit. The number of neurons in the hidden layer has been tested with k-fold cross validation and we have found $1000$ for the binary classification and $700$ for the multi-class case. We also have to set up the learning rate : the optimal value we have found is $3$ for the binary case and $2$ for the multi-class one. Larger is the value, faster could be the convergence but it may diverge. By default, the activation function used is the \textit{tanh} one. We didn't see so much different using the \textit{sigmoid}. Finally, the algorithm uses stochastic gradient descent using a momentum of $0.5$, which might help to converge faster.

@@

The idea of bootstrap aggregation is to average several decision trees. It will help to reduce the variance and avoid overfitting. It generates $m$ new training sets of the same size, by sampling uniformly and with replacement the original training set. Then we have $m$ decision trees which will grows independently with their own generated training sets. Finally, to predict the class it takes the average over all the trees. In order to find those parameters $m$ and the maximum depth of the tree, we have also used k-fold cross validation and the best values are $m = 200$ and maximum depth of $512$ for the binary classification and for the multi-class case $m = 500$ and maximum depth also of $512$. Having a high maximum depth might lead to overfit because we specialize to much the features split, but if it is too small, it might underfit because the features are not well split in order to identify classes. For the parameter $m$, it is better to have a higher number in order to have a large number of trees which can give a better accuracy of the prediction and reduce the variance.

@@

The idea of random forests is similar than with bagging in a sense that we train several trees and take the average predictions. Like bagging, each trees have trained on a subset of the dataset. However, there is another aspect in random forest that bagging doesn't have: for each split, instead of searching a split among the overall features, we search only among a random subset of features. The size of this subset is a parameter $m$ as well as the maximum depth tree and the number of trees, like in bagging. Like bagging, random forests are useful because they help reducing the variance and might avoid overfitting. But the more interesting ideas is to select only a subset of the features in the split process. Moreover, it is efficient against imbalanced dataset.

@@

The given Piotr matlab toolboox [2] contains a method \textit{forestTrain} which takes the defined parameters and \textit{forestApply} for the prediction.

@@

Adaboosting [4] is another way to average trees. If we observe more deeply the bagging approach, some trees might have a very high variance and some a very slow. Averaging all those trees decrease the variance and have a small bias. Boosting is the opposite : decrease the bias when averaging and have a small variance. In order to achieve this, each data sample has an assigned non-negative weight. After each iteration, if the data sample is misclassified, its weight is increased for the next iteration. So misclassified data samples will have more importance. The parameters are the same as in the bagging and behave in the same manner. The Piotr matlab toolboox [2] contains a method \textit{adaBoostTrain} and \textit{adaBoostApply} for the prediction. We have only used it for the binary classification, because the implementation is not adapted for more classes. However, we should have used the same technique as SVM : OVA. The parameters we obtain with k fold cross validation are $1024$ for the number of trees and $4$ for the maximum depth.

@@

In order to evaluate our different models, we have split our data in two sets of $80\%$ and $20\%$ for the training and the test sets. We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set. We learn only on the training set and then use the test set in order to estimate the error using the \textit{Balanced Error Rate}. This error compute the average percentage of misclassification per class. It allows us to avoid the case of the $90\%/10\%$, predict only the class of the $90\%$ and have a good error. We have also used k-fold cross validation to tune the parameters up of our different models. We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set). We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.

@@

This section describes our models for the binary classification as well as the multi-class classification. We have essentially put our effort on the CNN features. The reason is because with only the HOG features, we couldn't get an error lower than $20\%$ for binary and multi-class classification. We have tried to take the whole matrix, only a subset of features with $M = 1000$, different methods but nothing worked. With the CNN features, we didn't have this problem. We think that the reason is because those features are too noisy : they came directly form the images. Some of them are rotated, inverted and so, the gradient is also transformed, which might be confusing for the learning. The CNN features are already computed from a convolutional neural networks which gives the features it has seen on the image with a specific weight. Because each sample has approximately $5\%$ of the overall features, they are fewer and are more specific than the HOG features.

@@

[1]    IMM2012-06284, R. B. Palm, "Prediction as a candidate for learning deep hierarchical models of data", 2012.

@@

[3]    Pattern Recognition and Machine Learning, Christopher Bishop, p 569-570, 2007.

[4]    R. Appel, T. Fuchs, P. Dollár, P. Perona; "Quickly Boosting Decision Trees – Pruning Underachieving Features Early," ICML 2013.
\end{document}
